---
title: "MATH1005 Project 1 - Group 55"
author: "500557358; 500480757 "
subtitle: "Project X"
date: "University of Sydney | Math1005 | 2020"
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    theme: flatly
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
---

<br>

# Executive Summary
- The aim of this report is to ...
- The main discoveries are ...

<br>

# Full Report

## Domain Knowledge

## Initial Data Analysis (IDA)
### Data Quality - Origins of Data

The data set was found in a Federal Bureau of Investigation website from the criminal justice information services division. The data set contains the number of crimes committed (per year) and the rate (per 100 000 inhabitants) of different types of crime such as violent crime and robbery from the years 1997 to 2016 in the United States of America. 

This source is reliable as it is written/collected by a trustworthy government source. The results are valid as the Federal Bureau of Investigation is a government organization and thus is trustworthy and would collect accurate information. The information would also be checked and reviewed by others before being published on their website.

There are also limitations to this data set, this includes the missing data in the "rape" and "rape rate" variable. So that variable will not be not used in the analysis. However, missing data can lead to bias of the variables and results in a decrease in statistical power (Kang, 2013). 
Another limitation includes how the data set doesn't include crime that hasn't been reported to the police or FBI so thus the numbers in the data could be lower than the actual crimes committed in the U.S. This could lead to inaccurate data collected and therefore the conclusions or trends after analysis will be inaccurate. A research paper notes that crime statistics doesn't show the reality of crime, and can lead to an incorrect conclusion (Altbeker, 2016). Data also doesn't account for all types of crime including drug, cyber, financial crimes etc. 

## Complexity of Data
```{r}

crime = read.csv("crime.csv")

```
```{r}
#quick look at top 5 rows of data

head(crime)

```
```{r}

##Size of data 

dim(crime)

```
```{r}

## R's classification of data 

class(crime)

```
```{r}

## R's classification of variables

str(crime)

```
```{r}

## Reclassification of Variables to numerical

crime$Year=as.numeric(crime$Year)

```
Change Year variable from integer to numeric And change chr values to numeric. 
First have to get rid of the commas and spaces, then change them to numerical variables.

```{r}
## Remove commas and whitespaces
crime$Pop = gsub(",","",crime$Pop)
crime$Pop = gsub(" ","",crime$Pop)
crime$Pop = as.numeric(crime$Pop)

crime$ViolentCrime = gsub(",","",crime$ViolentCrime)
crime$ViolentCrime = gsub(" ","",crime$ViolentCrime)
crime$ViolentCrime = as.numeric(crime$ViolentCrime)

crime$MMS = gsub(",","",crime$MMS)
crime$MMS = gsub(" ","",crime$MMS)
crime$MMS = as.numeric(crime$MMS)

crime$Rape = gsub(",","",crime$Rape)
crime$Rape = gsub(" ","",crime$Rape)
crime$Rape = as.numeric(crime$Rape)

crime$Robbery = gsub(",","",crime$Robbery)
crime$Robbery = gsub(" ","",crime$Robbery)
crime$Robbery = as.numeric(crime$Robbery)

crime$AggAssault = gsub(",","",crime$AggAssault)
crime$AggAssault = gsub(" ","",crime$AggAssault)
crime$AggAssault = as.numeric(crime$AggAssault)

crime$PropCrime = gsub(",","",crime$PropCrime)
crime$PropCrime = gsub(" ","",crime$PropCrime)
crime$PropCrime = as.numeric(crime$PropCrime)

crime$Burg = gsub(",","",crime$Burg)
crime$Burg = gsub(" ","",crime$Burg)
crime$Burg = as.numeric(crime$Burg)

crime$LarcTheft = gsub(",","",crime$LarcTheft)
crime$LarcTheft = gsub(" ","",crime$LarcTheft)
crime$LarcTheft = as.numeric(crime$LarcTheft)

crime$LarcTheftRate = gsub(",","",crime$LarcTheftRate)
crime$LarcTheftRate = gsub(" ","",crime$LarcTheftRate)
crime$LarcTheftRate = as.numeric(crime$LarcTheftRate)

crime$VehicleTheftRate = gsub(",","",crime$VehicleTheftRate)
crime$VehicleTheftRate = gsub(" ","",crime$VehicleTheftRate)
crime$VehicleTheftRate = as.numeric(crime$VehicleTheftRate)

crime$VehicleTheft = gsub(",","",crime$VehicleTheft)
crime$VehicleTheft = gsub(" ","",crime$VehicleTheft)
crime$VehicleTheft = as.numeric(crime$VehicleTheft)
```
```{r}

## R's classification of variables after variable change

str(crime)

```
Summary:

- The data came from the Federal Bureau of Investigation, criminal justice information services division website.
- The data is valid because the FBI is an official government organization. 
- Possible issues include the missing data for rape and the data only having reported crimes.
- Each row represents the amount or rate (per 100 000) of a type of crime from the years 1997 to 2016.
- Each column represents the type (20 types) of crime committed, eg theft, rape etc.

<br>

## Research Question 1

Insert text and analysis. 

Summary: 

<br>

## Research Question 2

Insert text and analysis. 

Summary: 

<br>

# References (if needed)

Style: APA 

Altbeker, A. (2016). THE DANGERS OF DATA: Recognising the limitations of crime statistics. South African Crime Quarterly, 14. https://doi.org/10.17159/2413-3108/2005/v0i14a1004
  
  Kang, H., & Kang, H. (2013). The prevention and handling of the missing data. Korean Journal of Anesthesiology, 64(5), 402–406. https://doi.org/10.4097/kjae.2013.64.5.402


<br>
