---
title: "MATH1005 - Group 55"
author: "500557358; 500480757,500179286 "
subtitle: "Project 1"
date: "University of Sydney | Math1005 | 2020"
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    theme: flatly
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
---

<br>

# Executive Summary
- The aim of this report is to ...
- The main discoveries are ...

<br>

# Full Report

## Domain Knowledge

## Initial Data Analysis (IDA)
### Data Quality - Origins of Data

The data set was found in a Federal Bureau of Investigation website from the criminal justice information services division. The data set contains the number of crimes committed (per year) and the rate (per 100 000 inhabitants) of different types of crime such as violent crime and robbery from the years 1997 to 2016 in the United States of America. 

This source is reliable as it is written/collected by a trustworthy government source. The results are valid as the Federal Bureau of Investigation is a government organization and thus is trustworthy and would collect accurate information. The information would also be checked and reviewed by others before being published on their website.

There are also limitations to this data set, this includes the missing data in the "rape" and "rape rate" variable. So that variable will not be not used in the analysis. However, missing data can lead to bias of the variables and results in a decrease in statistical power (Kang, 2013). 
Another limitation includes how the data set doesn't include crime that hasn't been reported to the police or FBI so thus the numbers in the data could be lower than the actual crimes committed in the U.S. This could lead to inaccurate data collected and therefore the conclusions or trends after analysis will be inaccurate. A research paper notes that crime statistics doesn't show the reality of crime, and can lead to an incorrect conclusion (Altbeker, 2016). Data also doesn't account for all types of crime including drug, cyber, financial crimes etc. 

## Complexity of Data
```{r}

crime = read.csv("crime.csv")

```
```{r}
#quick look at top 5 rows of data

head(crime)

```
```{r}

##Size of data 

dim(crime)

```
```{r}

## R's classification of data 

class(crime)

```
```{r}

## R's classification of variables

str(crime)

```
```{r}

## Reclassification of Variables to numerical

crime$Year=as.numeric(crime$Year)

```
Change Year variable from integer to numeric And change chr values to numeric. 
First have to get rid of the commas and spaces, then change them to numerical variables.

```{r}
## Remove commas and whitespaces
crime$Pop = gsub(",","",crime$Pop)
crime$Pop = gsub(" ","",crime$Pop)
crime$Pop = as.numeric(crime$Pop)

crime$ViolentCrime = gsub(",","",crime$ViolentCrime)
crime$ViolentCrime = gsub(" ","",crime$ViolentCrime)
crime$ViolentCrime = as.numeric(crime$ViolentCrime)

crime$MMS = gsub(",","",crime$MMS)
crime$MMS = gsub(" ","",crime$MMS)
crime$MMS = as.numeric(crime$MMS)

crime$Rape = gsub(",","",crime$Rape)
crime$Rape = gsub(" ","",crime$Rape)
crime$Rape = as.numeric(crime$Rape)

crime$Robbery = gsub(",","",crime$Robbery)
crime$Robbery = gsub(" ","",crime$Robbery)
crime$Robbery = as.numeric(crime$Robbery)

crime$AggAssault = gsub(",","",crime$AggAssault)
crime$AggAssault = gsub(" ","",crime$AggAssault)
crime$AggAssault = as.numeric(crime$AggAssault)

crime$PropCrime = gsub(",","",crime$PropCrime)
crime$PropCrime = gsub(" ","",crime$PropCrime)
crime$PropCrime = as.numeric(crime$PropCrime)

crime$Burg = gsub(",","",crime$Burg)
crime$Burg = gsub(" ","",crime$Burg)
crime$Burg = as.numeric(crime$Burg)

crime$LarcTheft = gsub(",","",crime$LarcTheft)
crime$LarcTheft = gsub(" ","",crime$LarcTheft)
crime$LarcTheft = as.numeric(crime$LarcTheft)

crime$LarcTheftRate = gsub(",","",crime$LarcTheftRate)
crime$LarcTheftRate = gsub(" ","",crime$LarcTheftRate)
crime$LarcTheftRate = as.numeric(crime$LarcTheftRate)

crime$VehicleTheftRate = gsub(",","",crime$VehicleTheftRate)
crime$VehicleTheftRate = gsub(" ","",crime$VehicleTheftRate)
crime$VehicleTheftRate = as.numeric(crime$VehicleTheftRate)

crime$VehicleTheft = gsub(",","",crime$VehicleTheft)
crime$VehicleTheft = gsub(" ","",crime$VehicleTheft)
crime$VehicleTheft = as.numeric(crime$VehicleTheft)
```
```{r}

## R's classification of variables after variable change

str(crime)

```
Summary:

- The data came from the Federal Bureau of Investigation, criminal justice information services division website.
- The data is valid because the FBI is an official government organization. 
- Possible issues include the missing data for rape and the data only having reported crimes.
- Each row represents the amount or rate (per 100 000) of a type of crime from the years 1997 to 2016.
- Each column represents the type (20 types) of crime committed, eg theft, rape etc.

<br>

## Research Question 1

-Distribution of crime
 
```{r}
crime<-read.csv("crime.csv",header = TRUE);data
VCRate=data$VCRate
MMSRate=data$MMSRate
RapeRate=data$RapeRate
RobberyRate=data$RobberyRate
AggAssaultRate=data$AggAssault.1
BurgRate=data$BurgRate
LarcTheftRate=data$LarcTheftRate
VehicleTheftRate=data$VehicleTheftRate
boxplot(VCRate,horizontal = T)
#violent crime
boxplot(MMSRate,horizontal = T)
#MMS
boxplot(RapeRate,horizontal = T)
#Rape
boxplot(RobberyRate,horizontal = T)
#robbery
boxplot(AggAssaultRate,horizontal = T)
#agg assault
boxplot(BurgRate,horizontal = T)
#burg
boxplot(VehicleTheftRate,horizontal = T)
#vehicle thef
```
Summary: 
-According to FBI data, There are several types of crimes, including violent crime, MMS crime, rape, robbery, Aggravated Assault, burg, theft and vehicle theft. The three crimes with the largest number of crimes are Larc Theft, burg and violent crimes. Among them, the number of people who committed theft is the largest. This may be because the crime of theft is not severe, and theft and burglary are more profitable. If it does not involve overly expensive items, the criminals will usually not stay in prison for too long. A large number of violent crimes may be due to the broad classification, covering a wide range of crimes, leading to more criminals. The two relatively few items are MMS crime and rape. For the total number of crimes, the lower number of rapes is because it is a felony in the United States. Serious consequences and longer sentences have resulted in fewer crimes. 
-Looking at the data as a whole, from 1997 to 2016, the number of crimes showed a downward trend. Although not every year is less than the previous year, overall, the crime rate is declining. It can be seen from the box plot that although the crime rate is declining, the distribution of crime has not changed much. Comparing 1997 and 2016, burglary and theft are still the most common crimes, while MMS and rape are still the least common crimes. During the period 1997-2016, the distribution of crimes did not change significantly.


<br>

## Research Question 2

Insert text and analysis. 

Summary: 

<br>

# References (if needed)

Style: APA 

<br>


<br>

## Research Question 2

Insert text and analysis. 

Summary: 

<br>

# References (if needed)

Style: APA 

Altbeker, A. (2016). THE DANGERS OF DATA: Recognising the limitations of crime statistics. South African Crime Quarterly, 14. https://doi.org/10.17159/2413-3108/2005/v0i14a1004
  
  Kang, H., & Kang, H. (2013). The prevention and handling of the missing data. Korean Journal of Anesthesiology, 64(5), 402–406. https://doi.org/10.4097/kjae.2013.64.5.402


<br>
